\subsection{ dirtyfindzeros }
\noindent Name: \textbf{dirtyfindzeros}\\
gives a list of numerical values listing the zeros of a function on an interval.\\

\noindent Usage: 
\begin{center}
\textbf{dirtyfindzeros}(\emph{f},\emph{I}) : (\textsf{function}, \textsf{range}) $\rightarrow$ \textsf{list}\\
\end{center}
Parameters: 
\emph{f} is a function.\\
\emph{I} is an interval.\\

\noindent Description: \begin{itemize}

\item \textbf{dirtyfindzeros}(\emph{f},\emph{I}) returns a list containing some zeros of \emph{f} in the
   interval \emph{I}. The values in the list are numerical approximation of the exact
   zeros. The precision of these approximations is approximately the precision
   stored in \textbf{prec}. If \emph{f} does not have two zeros very close to each other, it 
   can be expected that all zeros are listed. However, some zeros may be
   forgotten. This command should be considered as a numerical algorithm and
   should not be used if safety is critical.

\item More precisely, the algorithm relies on global variables \textbf{prec} and \textbf{points} and
   is the following: let $n$ be the value of variable \textbf{points} and $t$ be the value
   of variable \textbf{prec}.
   \begin{itemize}
   \item  Evaluate $|f|$ at $n$ evenly distributed points in the interval $I$.
     the precision used is automatically chosen in order to ensure that the sign
     is correct.
   \item  Whenever $f$ changes its sign for two consecutive points,
     find an approximation $x$ of its zero with precision $t$ using
     Newton's algorithm. The number of steps in Newton's iteration depends on $t$:
     the precision of the approximation is supposed to be doubled at each step.
   \item  Add this value to the list.
   \end{itemize}
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{14.8cm}\begin{Verbatim}[frame=single]
   > dirtyfindzeros(sin(x),[-5;5]);
   [|-0.31415926535897932384626433832795028841971693993750801e1, 0, 0.31415926535897932384626433832795028841971693993750801e1|]
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{14.8cm}\begin{Verbatim}[frame=single]
   > L1=dirtyfindzeros(x^2*sin(1/x),[0;1]);
   > points=1000!;
   > L2=dirtyfindzeros(x^2*sin(1/x),[0;1]);
   > length(L1); length(L2);
   18
   25
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{prec}, \textbf{points}, \textbf{findzeros}
