\subsection{doubledouble}
\label{labdoubledouble}
\noindent Names: \textbf{doubledouble}, \textbf{DD}\\
represents a number as the sum of two IEEE doubles.\\

\noindent Description: \begin{itemize}

\item \textbf{doubledouble} is both a function and a constant.

\item As a function, it rounds its argument to the nearest number that can be written
   as the sum of two double precision numbers.

\item The algorithm used to compute \textbf{doubledouble}(x) is the following: let xh = \textbf{double}(x)
   and let xl = \textbf{double}(x-xh). Return the number xh+xl. Note that if the current 
   precision is not sufficient to represent exactly xh+xl, a rounding will occur
   and the result of \textbf{doubledouble}(x) will be useless.

\item As a constant, it symbolizes the double-double precision format. It is used in 
   contexts when a precision format is necessary, e.g. in the commands 
   \textbf{roundcoefficients} and \textbf{implementpoly}.
   See the corresponding help pages for examples.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity=1!;
> a = 1+ 2^(-100);
> DD(a);
0.10000000000000000000000000000007888609052210118054e1
> prec=50!;
> DD(a);
Warning: double rounding occurred on invoking the double-double rounding operato
r.
Try to increase the working precision.
1
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{double} (\ref{labdouble}), \textbf{doubleextended} (\ref{labdoubleextended}), \textbf{tripledouble} (\ref{labtripledouble}), \textbf{roundcoefficients} (\ref{labroundcoefficients}), \textbf{implementpoly} (\ref{labimplementpoly})
