\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> restart;
The tool has been restarted.
> prec=50;
The precision has been set to 50 bits.
> f=cos(2*exp(x));
> d=[-1/8;1/8];
> p=remez(f,2,d);
> derivativeZeros = dirtyfindzeros(diff(p-f),d);
> derivativeZeros = inf(d).:derivativeZeros:.sup(d);
> max=0;
Warning: syntax error, unexpected EQUALTOKEN, expecting LPARTOKEN.
The last symbol read has been "=".
Will skip input until next semicolon after the unexpected token. May leak memory
.
> for t in derivativeZeros do {
>    r = evaluate(abs(p-f), t);
>    if r > max then { max=r; argmax=t; };
Warning: syntax error, unexpected THENTOKEN, expecting LPARTOKEN.
The last symbol read has been "then".
Will skip input until next semicolon after the unexpected token. May leak memory
.
Warning: the identifier "t" is neither assigned to, nor bound to a library funct
ion nor external procedure, nor equal to the current free variable.
Will interpret "t" as "x".
Warning: syntax error, unexpected RIGHTCURLYBRACETOKEN.
The last symbol read has been "}".
Will skip input until next semicolon after the unexpected token. May leak memory
.
> };
Warning: syntax error, unexpected RIGHTCURLYBRACETOKEN.
The last symbol read has been "}".
Will skip input until next semicolon after the unexpected token. May leak memory
.
> print("The infinity norm of", p-f, "is", max, "and is reached at", argmax);
Warning: syntax error, unexpected COMMATOKEN, expecting LPARTOKEN.
The last symbol read has been ",".
Will skip input until next semicolon after the unexpected token. May leak memory
.
\end{Verbatim}
\end{minipage}\end{center}
