\subsection{remez}
\label{labremez}
\noindent Name: \textbf{remez}\\
computes the minimax of a function on an interval.\\
\noindent Usage: 
\begin{center}
\textbf{remez}(\emph{f}, \emph{n}, \emph{range}, \emph{w}, \emph{quality}) : (\textsf{function}, \textsf{integer}, \textsf{range}, \textsf{function}, \textsf{constant}) $\rightarrow$ \textsf{function}\\
\textbf{remez}(\emph{f}, \emph{L}, \emph{range}, \emph{w}, \emph{quality}) : (\textsf{function}, \textsf{list}, \textsf{range}, \textsf{function}, \textsf{constant}) $\rightarrow$ \textsf{function}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{f} is the function to be approximated
\item \emph{n} is the degree of the polynomial that must approximate \emph{f}
\item \emph{L} is a list of monomials that can be used to represent the polynomial that must approximate \emph{f}
\item \emph{range} is the interval where the function must be approximated
\item \emph{w} (optional) is a weight function. Default is 1.
\item \emph{quality} (optional) is a parameter that controls the quality of the returned polynomial \emph{p}, with respect to the exact minimax $p^\star$. Default is 1e-5.
\end{itemize}
\noindent Description: \begin{itemize}

\item \\textbf{remez} computes an approximation of the function $f$ with respect to \n   the weight function $w$ on the interval \\emph{range}. More precisely, it \n   searches a polynomial $p$ such that $\\|pw-f\\|_{\\infty}$ is \n   (almost) minimal among all polynomials $p$ of a certain form. The norm is\n   the infinity norm, e.g. $\\|g\\|_{\\infty} = \\max \\{|g(x)|, x \\in \\mathrm{range}\\}.$\n
\item If $w=1$ (the default case), it consists in searching the best \n   polynomial approximation of $f$ with respect to the absolute error.\n   If $f=1$ and $w$ is of the form $1/g$, it consists in \n   searching the best polynomial approximation of $g$ with respect to the \n   relative error.\n
\item If $n$ is given, the polynomial $p$ is searched among the \n   polynomials with degree not greater than $n$.\n   If \\emph{L} is given, the polynomial $p$ is searched as a linear combination \n   of monomials $X^k$ where $k$ belongs to \\emph{L}.\n   \\emph{L} may contain ellipses but cannot be end-elliptic.\n
\item The polynomial is obtained by a convergent iteration called Remez' algorithm. \n   The algorithm computes a sequence $p_1,\\dots ,p_k,\\dots$ \n   such that $e_k = \\|p_k w-f\\|_{\\infty}$ converges towards \n   the optimal value $e$. The algorithm is stopped when the relative error \n   between $e_k$ and $e$ is less than \\emph{quality}.\n
\item Note: the algorithm may not converge in certain cases. Moreover, it may \n   converge towards a polynomial that is not optimal. These cases correspond to \n   the cases when the Haar condition is not fulfilled.\n   See [Cheney - Approximation theory] for details.\n\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 3: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{dirtyinfnorm} (\ref{labdirtyinfnorm}), \textbf{infnorm} (\ref{labinfnorm})
