\subsection{unsuppressmessage}
\label{labunsuppressmessage}
\noindent Name: \textbf{unsuppressmessage}\\
\phantom{aaa}unsuppresses the displaying of messages with a certain number\\[0.2cm]
\noindent Library name:\\
\verb|   void sollya_lib_unsuppressmessage(sollya_obj_t, ...);|\\[0.2cm]
\noindent Usage: 
\begin{center}
\textbf{unsuppressmessage}(\emph{msg num 1}, ..., \emph{msg num n})   : (\textsf{integer}, ..., \textsf{integer}) $\rightarrow$ \textsf{void}\\
\textbf{unsuppressmessage}(\emph{msg list})   : \textsf{list} $\rightarrow$ \textsf{void}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{msg num 1} thru \emph{msg num n} represent the numbers of $n$ messages to be suppressed
\item \emph{msg list} represents a list with numbers of messages to be suppressed
\end{itemize}
\noindent Description: \begin{itemize}

\item The \textbf{unsuppressmessage} command allows particular warning and information
   messages that have been suppressed from message output to be
   unsuppressed, i.e. activated for display again. Every \sollya warning
   or information message (that is not fatal to the tool's execution) has
   a message number. When these message numbers \emph{msg num 1} thru \emph{msg num n} 
   are given to \textbf{unsuppressmessage}, the corresponding message are displayed
   again, as they are by default at according verbosity levels. Actually,
   the \textbf{unsuppressmessage} command just reverts the effects of the \textbf{suppressmessage}
   command.

\item Instead of giving \textbf{unsuppressmessage} several message numbers \emph{msg num 1} thru
   \emph{msg num n} or calling \textbf{unsuppressmessage} several times, it is possible to give
   a whole list \emph{msg list} of message numbers to \textbf{unsuppressmessage}.

\item The user should be aware that \textbf{unsuppressmessage} presents sticky behavior for
   the warning and information messages suppressed from output.  In fact,
   \textbf{unsuppressmessage} just unsuppresses the warning or information messages given
   in argument.  All other suppressed messages stay suppressed until they
   get unsuppressed by subsequent calls to \textbf{unsuppressmessage} or the \sollya tool
   is restarted. This behavior distinguishes message suppression from
   other global states of the \sollya tool. The user may use
   \textbf{getsuppressedmessages} to obtain a list of currently suppressed
   messages. In particular, in order to unsuppressed all currently
   suppressed warning or information messages, the user may feed the
   output of \textbf{getsuppressedmessages} (a list) into \textbf{unsuppressmessage}.

\item The user should also note that unsuppressing warning or information
   messages with \textbf{unsuppressmessage} just reverts the effects of the \textbf{suppressmessage}
   command but that other conditions exist that affect the actual displaying 
   of a message, such as global verbosity (see \textbf{verbosity}) and modes
   like rounding warnings (see \textbf{roundingwarnings}). A message will not just 
   get displayed because it was unsuppressed with \textbf{unsuppressmessage}.

\item When \textbf{unsuppressmessage} is used on message numbers that do not exist in the
   current version of the tool, a warning is displayed. The call has no
   other effect though.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity = 1;
The verbosity level has been set to 1.
> 0.1;
Warning: Rounding occurred when converting the constant "0.1" to floating-point 
with 165 bits.
If safe computation is needed, try to increase the precision.
0.1
> suppressmessage(174);
> 0.1;
0.1
> suppressmessage(174);
> 0.1;
0.1
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity = 12;
The verbosity level has been set to 12.
> showmessagenumbers = on; 
Displaying of message numbers has been activated.
> diff(exp(x * 0.1));
Warning (174): Rounding occurred when converting the constant "0.1" to floating-
point with 165 bits.
If safe computation is needed, try to increase the precision.
Information (196): formally differentiating a function.
Information (197): differentiating the expression 'exp(x * 0.1)'
Information (204): extraction of coefficient terms from a polynomial uses a spec
ial algorithm for Horner forms.
Information (207): no Horner simplification will be performed because the given 
tree is already in Horner form.
exp(x * 0.1) * 0.1
> suppressmessage([| 174, 207, 196 |]);
> diff(exp(x * 0.1));
Information (197): differentiating the expression 'exp(x * 0.1)'
Information (204): extraction of coefficient terms from a polynomial uses a spec
ial algorithm for Horner forms.
exp(x * 0.1) * 0.1
> unsuppressmessage([| 174, 196 |]);
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 3: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity = 12;
The verbosity level has been set to 12.
> showmessagenumbers = on;
Displaying of message numbers has been activated.
> suppressmessage(207, 387, 390, 388, 391, 196, 195, 197, 205);
> getsuppressedmessages();
[|195, 196, 197, 205, 207, 387, 388, 390, 391|]
> evaluate(x/sin(x) - 1, [-1;1]);
[0;0.85081571768092561791175324139865019347039665509409]
> unsuppressmessage(getsuppressedmessages());
> getsuppressedmessages();
[| |]
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{getsuppressedmessages} (\ref{labgetsuppressedmessages}), \textbf{suppressmessage} (\ref{labsuppressmessage}), \textbf{unsuppressmessage} (\ref{labunsuppressmessage}), \textbf{verbosity} (\ref{labverbosity}), \textbf{roundingwarnings} (\ref{labroundingwarnings})
